Our Title

### 2015-05-06 draft.tex
Added in James B's "Dissenting Opinion" chapter.

\chapter{???????   Things we still need to hit   ?????????}

>>>>

\chapter{Dissention Minority Opinion (Or Option B or Some other name)}

\begin{center}
 \enquote{One of the most important lessons of the Cold War was that incessantly worrying about the low-probability/high-impact cases was a mistake.}  - Kenneth Pollack \cite{pollack2014unthinkable}
\end{center}

The ISIL nuclear threat is, by definition, a low-probability/high-impact case as this report has shown. As Pollack points out, historically the low-probability/high-impact events receive a disproportionate amount of the national attention and funding \cite{pollack2014unthinkable}. Americans in particular are susceptible to the allure of these \enquote{grand challenges} and often pour massive amounts of monetary and human resources into developing solutions, but the U.S. is certainly not alone in this regard. Staying within the nuclear realm, both the U.S. and U.S.S.R. both devoted significant attention to the \enquote{bolt from the blue} problem during the Cold War, despite there never being evidence that such a tactic was considered or likely to succeed \cite{pollack2014unthinkable,Bracken2013}.  In fact, some of this planning still takes placed even today, despite 20+ years passing since the end of the Cold War.

In addition to the inefficient use of government resources, the focus on rare events can lead to the misinterpretation of signals and an exacerbation of the original dominant security risk.  For example, the U.S.S.R interpreted the 1983 Able Archer NATO nuclear exercise as being a preparation for a \enquote{bolt from the blue} attack \cite{pollack2014unthinkable}. As applied to ISIL, the remote threat of nuclear terrorism can serve as a distractor from the actual security risks posed by ISIL to U.S. interests in the region.  Additionally, it would result in less resources and political capital to devote to addressing the economic, political, and human rights issues that gives rise to ISIL and similar movements throughout the Middle East in the first place \cite{Morand2015}. Unintentional second-order consequences could result in the U.S. and international focus on nuclear terrorism issue creating the environment, public awareness and making a nuclear capable ISIL credible, for ISIL to achieve its goals through propaganda without actually developing nuclear capabilities.  For example, while not well orchestrated by ISIL standards and reported to an unbelieving public, there have been previous overstated reports of ISIL's nuclear capabilities \cite{Bunn2014}.  In the future, it may be hard to tell fact from fiction or to prove a negative to a skeptical American public.  While this strategy could backfire on ISIL, most notably through the deployment of U.S. forces to combat the threat, a more likely result under the current Administration is that the U.S. would seek diplomatic and technical solutions.

In fact, diplomatic and technical solutions are the bulk of the policy proposals put forth in this report.  In theory, this is a great approach as it would not be difficult to influence the American people or U.S. government to support efforts to combat terrorism, nuclear or non-nuclear, in the post-9/11 environment. Since 9/11, the U.S. has largely adopted a \enquote{1\% Doctrine} where all threats with even a 1\% likelihood should be eliminated \cite{pollack2014unthinkable}. This, coupled with a general reluctance of the American populace to question the U.S. military or anything posed as a national security issue, has given de facto carte blanche to all \enquote{good idea fairies} to promote their particular product or idea in the name of improving national security \cite{Fallows2015}. A relevant example to the proposed policies in this report is the history and effectiveness of radiation portal monitors (RPMs) and the Advance Spectroscopic Portals (ASPs).

RPMS evolved out of a layered defense concept to prevent a low-probability/high-impact scenario, nuclear terrorism. This layered defense is composed of RPMs as discussed in

\section{??????? Insert cross-ref here ??????}

Section XXX INSERT REF TO RPM section here XXX, the overseas equivalent known as the SLD as discussed in

\section{??????? Insert cross-ref here ??????}

Section XXX INSERT REF TO SLD SECTION HERE XXX, and the Global Threat Reduction Initiative discussed in

\section{??????? Insert cross-ref here ??????}

Section XXX INSERT REF TO NUCLEAR THREAT REDUCTION INITIATIVE SECTION HERE XXX.  Despite the layered technical and diplomatic approach, there are significant gaps in the ability to intercept a shipment into the U.S. containing SNM.  For example, DHS has implemented 99\% screening at seaports, but struggles to implement a significant screening at airports, due to a lack of natural choke points and disruption caused by screenings, and at the U.S.'s notoriously porous border crossings \cite{DepartmentofHomelandSecurityDHS2013,GovernmentAccountabilityOffice2012}.  Even if a shipment does cross a border, seaport, or airport with a RPM, the detection probability is only 90\% in the best of scenarios, if the RPM is even operational at the time \cite{DepartmentofHomelandSecurityDHS2013,DepartmentofHomelandSecurityDHS2006}.  To top of the list of issues, the ANSI standard for certification of RPMs contains no SNM requirements due to the inherent difficult in passive detection of SNM \cite{Ansi2007}.  The ASP program was conceived to address some of these problems and increase interception through advanced detectors and algorithms, but it failed to pass 6 of 10 criteria in field tests, criteria which were widely criticized as being favorable to the vendors in the first place \cite{UnitedStatesGovernmentAccountabilityOffice2013}. To further complication matters, if a competent proliferator uses shielding or naturally occurring radioactive material within the shipment, it is effectively impossible to pull out the SNM signal with the passive detection systems currently employed without generating significant false positives \cite{UnitedStatesGovernmentAccountabilityOffice2013,UnitedStatesCongress.House.CommitteeonHomelandSecurity.SubcommitteeonEmergingThreatsCybersecurity2007}.  Nonetheless, DNDO continues to invest significant effort into a technical solution that is at best 90\% effective, and at worst almost useless, with the goal of deterring an extremely low probability event.  The proposed policies in this report follow that same trend and fail to learn the lessons of history in this regard.

While avoiding unintended consequences or wasted resources are compelling reasons to reconsider the proposed policies, perhaps the most compelling reason is to place the threat and proposed policies into context.  It is perhaps a bit generous to even call ISIL obtaining a nuclear weapon a low probability event as there is zero basis in history for it to be able to do so \cite{pollack2014unthinkable,Bracken2013,Reed2010,Agency2004}. Developing nuclear weapons amidst crippling sanctions, airstrikes, and a dedicated opposition has never been done, despite several nation states attempting to do so \cite{Reed2010,Agency2004}.  In fact, the only nation to develop a NW in the time of war was the U.S., and the U.S. mainland was never seriously threatened. Supposing that ISIL was the proverbial \enquote{Black Swan} case, it is worth noting that, with the exception of the U.S., no state has gone from decision to having a deliverable nuclear weapon in under four years \cite{Reed2010}. Even modern conservative estimates place the timeline at 5-6 years for a competent nation state pursuing an optimal path to a nuclear weapon \cite{Harney2006}. Proponents of the threat of nuclear terrorism will likely point to the possibility that ISIL will circumvent the indigenous development route completely and pursue theft or purchase of key components or a complete device.  Again, there is no history to suggest that a nation state would take that risk given the enormous repercussions it would face \cite{pollack2014unthinkable}.  Additionally, anything short of a complete device would still require significant development to weaponized, perhaps three to four years depending on the amount of assistance ISIL received \cite{Harney2006}.

The sum result of this discussion is that we are at best facing the prospect of a nuclear armed ISIL in several years assuming no disruptions due to discovery and outside assistance from a nuclear capable state, both of which are counter to the history of every nuclear program \cite{pollack2014unthinkable,Reed2010}.  Given this fact, the best policy we have is to place financial, diplomatic, and military pressure on ISIL, all of which are current policies of the Obama Administration as shown in

\section{??????? Insert cross-ref here ??????}

Table XXX POLICY CHART XXX.  Additionally, there are no identified potential proliferation pathways that are not covered to some extent by current policies, programs, or operations.  As war was building in Europe in 1940, Edward Murrow reported, \enquote{The people here feel the machine is out of control, that we are all passengers on an express train traveling at high speed through a dark tunnel toward an unknown destiny.  The suspicion recurs that the train may have no engineer.}  The proposed policies are emblematic of a machine out of control in support of Dick Cheney's \enquote{1\% Doctrine.}  Therefore, it is recommended to not repeat the mistakes of history. Instead, it is proposed the U.S. \enquote{stays the course} and makes no changes to U.S. policy specifically to prevent the low probability event that is ISIL obtaining nuclear weapons.
